\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-04-09 ase20/parser.tex
[ASE] Added references / Changed red parts to black

\textbf{Background: Parsing Expression Grammar.}
Most parser generators target context-free languages with specific parsing
algorithms for Context-Free Grammar (CFG): JavaCC with LL(k)~\inred{\cite{llk}},
Bison with GLR~\inred{\cite{glr}}, and ANTLR with
ALL(*)~\inred{\cite{all-star}}.  However, they are not directly applicable for the
ECMAScript syntax because ECMAScript lexical and syntactic grammars require
context-sensitive lexers and parsers:
\begin{itemize}%[leftmargin=0.5cm]
  \item \textbf{Context-sensitive tokens:}
    ECMAScript tokens are context-sensitive because of JavaScript regular
    expressions and template strings.  For example, \( \code{/x/g} \) could be a
    single regular expression token or four tokens that represent division by
    variables \( \code{x} \) and \( \code{g} \) depending on enclosing contexts.
    Thus, lexers should be evaluated during parsing not before parsing.
  \item \textbf{Context-sensitive \( \bnfes \) symbols:}
    \( \bnfes \) supports context-sensitive symbols, which are positive (negative)
    lookahead \( +\symb \) (\( -\symb \)), exclusion \( \symb \butnot \symb' \),
    and no line-terminator \( \nolt \).  They are highly expressive
    and they can even represent the classic non-context-free language \( \{ a^n b^n c^n
    : n \geq 1 \} \) with the following productions:
    \[
      \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
        \NT{S} &::=& +(\NT{X}~\T{c})~\NT{A}~\NT{Y}&
        \NT{X} &::=& \T{a}~\NT{X}?~\T{b}\\
        \NT{A} &::=& \T{a}~\NT{A}?&
        \NT{Y} &::=& \T{b}~\NT{Y}?~\T{c}\\
      \end{array}
    \]
    However, it is not trivial to support such \( \bnfes \) symbols in CFG-based
    parser generators.
\end{itemize}

@@

\textbf{Parser Generation.}
The next step is to automatically synthesizes parsers for each production in \(
\bnfes \).  We extended Scala parser combinators to support lookahead parsing
and \( \bnfes \) notations. For example, the synthesized parser from the
production \textit{ArrayLiteral} in Figure~\ref{fig:array-literal}(a) is the one
in Figure~\ref{fig:array-literal}(b).  A na\"{i}ve implementation of lookahead
parsing would take exponential time because of backtracking.  To reduce it
to linear time, we applied the memoization technique introduced in
Packrat parsing~\cite{packrat}.  \inred{Moreover, we also
implemented the \textit{growing the seed} technique presented
by~\citet{packrat-lr} to support direct and even indirect left recursive
productions.  It enables the synthesis of parsers without changing the structure of
each production in \( \bnfes \).}

>>>>

\textbf{Background: Parsing Expression Grammar.}
Most parser generators target context-free languages with specific parsing
algorithms for Context-Free Grammar (CFG): JavaCC with LL(k)~\cite{llk}, Bison
with GLR~\cite{glr}, and ANTLR with ALL(*)~\cite{all-star}.  However, they are
not directly applicable for the ECMAScript syntax because ECMAScript lexical and
syntactic grammars require context-sensitive lexers and parsers:
\begin{itemize}%[leftmargin=0.5cm]
  \item \textbf{Context-sensitive tokens:}
    ECMAScript tokens are context-sensitive because of JavaScript regular
    expressions and template strings.  For example, \( \code{/x/g} \) could be a
    single regular expression token or four tokens that represent division by
    variables \( \code{x} \) and \( \code{g} \) depending on enclosing contexts.
    Thus, lexers should be evaluated during parsing not before parsing.
  \item \textbf{Context-sensitive \( \bnfes \) symbols:}
    \( \bnfes \) supports context-sensitive symbols, which are positive (negative)
    lookahead \( +\symb \) (\( -\symb \)), exclusion \( \symb \butnot \symb' \),
    and no line-terminator \( \nolt \).  They are highly expressive
    and they can even represent the classic non-context-free language \( \{ a^n b^n c^n
    : n \geq 1 \} \) with the following productions:
    \[
      \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
        \NT{S} &::=& +(\NT{X}~\T{c})~\NT{A}~\NT{Y}&
        \NT{X} &::=& \T{a}~\NT{X}?~\T{b}\\
        \NT{A} &::=& \T{a}~\NT{A}?&
        \NT{Y} &::=& \T{b}~\NT{Y}?~\T{c}\\
      \end{array}
    \]
    However, it is not trivial to support such \( \bnfes \) symbols in CFG-based
    parser generators.
\end{itemize}

@@

\textbf{Parser Generation.}
The next step is to automatically synthesizes parsers for each production in \(
\bnfes \).  We extended Scala parser combinators to support lookahead parsing
and \( \bnfes \) notations. For example, the synthesized parser from the
production \textit{ArrayLiteral} in Figure~\ref{fig:array-literal}(a) is the one
in Figure~\ref{fig:array-literal}(b).  A na\"{i}ve implementation of lookahead
parsing would take exponential time because of backtracking.  To reduce it
to linear time, we applied the memoization technique introduced in
Packrat parsing~\cite{packrat}.  Moreover, we also implemented the
\textit{growing the seed} technique presented by~\citet{packrat-lr} to support
direct and even indirect left recursive productions.  It enables the synthesis
of parsers without changing the structure of each production in \( \bnfes \).
